Dealing with Consumer Uncertainty: Online Product Reviews and What makes them helpful?

### 2022-09-13 Paper.Rmd
extend nbt and loss aversion

-   expand NBT a little bit more

Negativity Bias Theorem argues that humans or even animals have tendencies to be more aware of negative entities than positive ones. Someone else's hair in one's meal has the potential of ruining the whole meal. On the contrary a plate of hair won't turn to a delicious meal just because there is a slice of a delicious pizza on top of the hair. @rozin2001 explains the Negativity Bias Theorem by describing it with four main aspects. Namely, Negative Potency, Steeper Negative Gradients, Negative Dominance and Negative differentiation [@rozin2001; @sen2007; @eslami2018].

**Negative Potency** defends that the exact opposite of something positive carries heavier weight. In other words, negative events are much more impactful than their exact positive counterparts. For instance, losing 100\$ could have a harder psychological effect on a person than that particular person is gaining 100\$. Loss Aversion [@kahneman1979; @tversky1991] is the phenomenon where the Negative Potency can be observed the best. People risk more/less when they are losing/gaining something. The behavior can be explained by the Endowment Effect [@kahneman1990]. Endowment effect occurs because people usually over-estimate the value of their belongings, for example a pencil. However, they would have under-valued the same pencil if the item didn't belong to them. The reason to this behavior is that the dis-utility of losing that pencil is greater than the utility of acquring the same pencil.

@@

**Negativity Dominance** asserts that a balanced mixture of positive and negative evaluations creates a perception of negativity. This means that negative entities have higher weight than positive entities. This is apparent in religions. For most of the religions being a good and pious person takes time, demands upmost care. However, tainting this image is very easy with certain actions such as murder or pre-martial intercourse.

@@

-   talk about loss aversion in detail

Prospect Theory (Loss Aversion) criticizes Expected Utility Theory by introducing the thought that Expected Utility Theory is insufficient explaining the decision making behavior of economic agents who are under risk. In Expected Utility Theory outcomes are weighted by their probabilities but Prospect Theory discusses that people give more weight to the outcomes that they know for sure will happen. This is called *certainty effect* [@kahneman1979].

Humans tend to care about losses more than same amount of gains. Losing 100\$ is psychologically much more impactful than finding 100\$. This behavior is best explained in "loss aversion" from behavioral economics. [@kahneman1979; @schindler2017]. Loss Aversion has a strong connection to Negativity Bias Theorem since negative cues might be used as an early warning against potential losses [@herr1991].

>>>>

Negativity Bias Theorem argues that humans or even animals have tendencies to be more aware of negative entities than positive ones. Someone else's hair in one's meal has the potential of ruining the whole meal. On the contrary a plate of hair won't turn to a delicious meal just because there is a slice of a delicious pizza on top of the hair. @rozin2001 explains the Negativity Bias Theorem by describing it with four main aspects. Namely, Negative Potency, Steeper Negative Gradients, Negative Dominance and Negative Differentiation [@rozin2001; @sen2007; @eslami2018]. Loss Aversion has a strong connection to Negativity Bias Theorem since negative cues might be used as an early warning against potential losses [@herr1991].

**Negative Potency** defends that the exact opposite of something positive carries heavier weight. In other words, negative events are much more impactful than their exact positive counterparts. For instance, losing 100\$ could have a harder psychological effect on a person than that particular person is gaining 100\$. Loss Aversion [@kahneman1979; @tversky1991] is the phenomenon where the Negative Potency can be observed the best. People risk more/less when they are losing/gaining something. This behavior can be explained by the Endowment Effect [@kahneman1990]. Endowment effect occurs because people usually over-estimate the value of their belongings, for example a pencil. However, they would have under-valued the same pencil if the item didn't belong to them. The reason to this behavior is that the dis-utility of losing that pencil is greater than the utility of acquiring the same pencil.

@@

**Negativity Dominance** asserts that a balanced mixture of positive and negative evaluations creates a perception of negativity. This means that negative entities have higher weight than positive entities. This is apparent in religions. For most of the religions being a good and pious person takes time, demands utmost care. However, tainting this image is very easy with certain actions such as murder or pre-martial intercourse.

@@

Prospect Theory (Loss Aversion) criticizes Expected Utility Theory by introducing the thought that Expected Utility Theory is insufficient explaining the decision making behavior of economic agents who are under risk. In Expected Utility Theory outcomes are weighted by their probabilities but Prospect Theory discusses that people dynamically give more or less weight to the outcomes depending on if it is a loss or a gain. This is called *certainty effect* [@kahneman1979]. The Certainty Effect causes people to behave differently whether they are going to lose or gain something. In the positive domain, people will prefer a small amount of a certain gain over a big gain which is merely probable (risk averse). On the contrary, in the negative domain, people will have a preference towards a big loss which is merely probable over a small certain loss (risk seeking).

Since the cognitive abilities of humans are limited they tend to focus on what is crucial to them. Therefore, in the information processing context and because of the existence of the negativity bias, it could be expected that people process negative cues more thoroughly than positive ones. Furthermore, people may develop much more advanced memory about negative entities. This can also cause high awareness to negative information [@baumeister2001; @taylor1991].
